---
title: "Analysis of German publication output"
author: "Anne Hobert"
date: "3/6/2020"
output:
  github_document:
  word_document:
    fig_caption: yes
urlcolor: blue
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "90%",
  fig.align = "center",
  dpi = 300
)
```

```{r}
library(tidyverse)
library(urltools)
library(cowplot)
library(viridis)
```

## Data

In this document we describe the analysis of our sample of publications from German research institutions. We work with the dataset `pubs_cat` generated in [data_gathering.Rmd](data_gathering.Rmd).

```{r}
pubs_cat <- readr::read_csv("data/pubs_cat.csv", col_types = "dcdcclcc")
```


## Investigation of research questions

The goal is to answer the three research questions 

1) Has the OA fraction of the publication output of German universities and research institutions increased constantly over time?
2) Can we observe differences between the research sectors of the German science system? Are there obvious explanations for this (like different missions or subject profiles?
3) Which OA type is the most prevalent OA approach and can we identify different patterns of adoption to OA?

### OA fraction of the German publication output

 - plot of oa numbers on national level (Is OA? plot) over time
 - possibly scatterplot with all insitutions, publication output vs. oa share
 
 First, we look at how the overall OA share developed over time. The following figure displays the number of publications associated with one of the German research institutions we considered and highlights they part that is freely accessible online according to Unpaywall over the considered time period from 2010 until 2018.
 
```{r, fig.cap="Open access to journal articles from German research institutions according to Unpaywall. Blue area represents journal articles with at least one freely available full-text, grey area represents toll-access articles."}
pubs_cat %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    is_oa = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  ))  %>%
  group_by(PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_area(aes(fill = fct_rev(oa_category), group = fct_rev(oa_category)),  alpha = 0.8) +
    scale_fill_manual(values = c("#b3b3b3a0", "#56B4E9"), name = "Is OA?", labels = c("Not OA", "OA"))+
    labs(x = "Publication Year", y = "Number of Articles", title = "Total number of OA Articles") +
    theme(legend.position = "right")
```

As can be seen, the total number of articles, as well as the part that is OA increases constantly over time. The number of articles that are not openly available, is quite stable with a slow increase from `r pubs_cat %>% filter(oa_category == "not_oa") %>% group_by(PUBYEAR) %>% summarise(n = n_distinct(PK_ITEMS)) %>% filter(PUBYEAR == 2010) %>% .$n` in 2010 to `r pubs_cat %>% filter(oa_category == "not_oa") %>% group_by(PUBYEAR) %>% summarise(n = n_distinct(PK_ITEMS)) %>% filter(PUBYEAR == 2013) %>% .$n` in 2013, and decreasing again continuously from that point onwards to `r pubs_cat %>% filter(oa_category == "not_oa") %>% group_by(PUBYEAR) %>% summarise(n = n_distinct(PK_ITEMS)) %>% filter(PUBYEAR == 2018) %>% .$n`  publications in 2018. Since the number of OA articles increases continuously from `r pubs_cat %>% filter(oa_category != "not_oa") %>% group_by(PUBYEAR) %>% summarise(n = n_distinct(PK_ITEMS)) %>% filter(PUBYEAR == 2010) %>% .$n` publications in 2010 to `r pubs_cat %>% filter(oa_category != "not_oa") %>% group_by(PUBYEAR) %>% summarise(n = n_distinct(PK_ITEMS)) %>% filter(PUBYEAR == 2018) %>% .$n` in 2018, the relative proportion of OA articles rises significantly from `r round(pubs_cat %>% filter(oa_category != "not_oa") %>% group_by(PUBYEAR) %>% summarise(n = n_distinct(PK_ITEMS)) %>% filter(PUBYEAR == 2010) %>% .$n/pubs_cat %>% group_by(PUBYEAR) %>% summarise(n = n_distinct(PK_ITEMS)) %>% filter(PUBYEAR == 2010) %>% .$n, 4)*100` % in 2010 to `r round(pubs_cat %>% filter(oa_category != "not_oa") %>% group_by(PUBYEAR) %>% summarise(n = n_distinct(PK_ITEMS)) %>% filter(PUBYEAR == 2018) %>% .$n/pubs_cat %>% group_by(PUBYEAR) %>% summarise(n = n_distinct(PK_ITEMS)) %>% filter(PUBYEAR == 2018) %>% .$n, 4)*100` % in 2018.

```{r}
pubs_cat %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    is_oa = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  ))  %>%
  group_by(PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  mutate(oa_category = ifelse(oa_category == "not_oa", "Not OA", "OA")) %>% 
  knitr::kable(
    col.names = c(
      "Publication Year",
      "Is OA ?",
      "Number of Articles"
    ),
    big.mark = ",",
    caption = "Number of articles that are openly accessible (OA) or not (Not OA) per year."
  )
```

### Differences between sectors

In order to investigate what role the different sectors play in OA publishing in Germany and how they contribute to the OA development/overall OA shares, we distplay the development over time of the number of OA articles for each sector in the following figure. Note that scales for the `y-axes` are not the same, since the total publication output varies significantly among sectors.

```{r, fig.asp=1, fig.cap="Open access to journal articles per sector according to Unpaywall. Blue area represents journal articles with at least one freely available full-text, grey area represents toll-access articles. Sectors are ordered by publication output with the highest output top left and lowest at the bottom."}
pubs_cat %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    is_oa = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  ))  %>%
  group_by(sector) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  group_by(sector, PUBYEAR, oa_category, n_total) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_area(aes(fill = fct_rev(oa_category), group = fct_rev(oa_category)),  alpha = 0.8) +
    scale_fill_manual(values = c("#b3b3b3a0", "#56B4E9"), name = "Is OA?", labels = c("Not OA", "OA"))+
    facet_wrap(~ fct_rev(fct_reorder(sector, n_total)), scales = "free", ncol = 2) +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of OA Articles per sector") +
    theme(legend.position = "right")
```
some text to the second figure here.

In order to investigate the variability of OA publishing within the sectors, we now go one level deeper and examine OA shares of individual institutions, grouped by the sector they belong to. To this end, we first calculate the individual OA shares.

```{r}
oa_shares <- pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    is_oa = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  )) %>% 
  group_by(INST_NAME) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>%
  group_by(INST_NAME, oa_category, n_total) %>% 
  summarise(n_cat = n_distinct(PK_ITEMS)) %>% 
  pivot_wider(names_from = oa_category,
              values_from = n_cat,
              values_fill = list(n_cat = 0)) %>% 
  mutate(oa_share = is_oa/n_total *100) %>%
  rename(n_oa = is_oa, n_not_oa = not_oa) %>% 
  ungroup()
```
The following figure displays scatterplots where the OA share of an institution over the whole time period is shown with respect to its publication output.

```{r, fig.asp=1, fig.cap="Open Access shares of research institutions in Germany with respect to their total publication output grouped by the sector they belong to."}
pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  )) %>% 
  group_by(sector) %>% 
  mutate(n_total_sec = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>%
  left_join(oa_shares, by = 'INST_NAME') %>% 
  group_by(sector, INST_NAME, n_total, oa_share, n_total_sec) %>%
  summarise() %>% 
  ggplot(aes(x = n_total, y = oa_share)) +
    geom_point(aes(color = sector)) + 
    scale_x_log10() + 
    scale_color_viridis(end = 0.95, direction = -1, option = "plasma", discrete = TRUE, name = "Sectors")+
    facet_wrap(~ fct_rev(fct_reorder(sector, n_total_sec)), ncol = 2) +
    labs(x = "Total publication output (logarithmic scale)", y = "OA share (in %)", title = "Open Access shares of research institutions in Germany") +
    theme(legend.position = "bottom")
```
The most striking observations from this figure are the high OA shares of most of the Max-Planck-institutes and the very low OA fractions of almost all of the state and federal institutes as well as the ones from the Fraunhofer Society. Universities and Leibniz-Society have many institutes with OA shares close to one half. We can further see very well that the universities have by far the largest publication volumes, followed by the Helmholtz-Society.

The following box plot quantifies the observations regarding the variability of OA shares within sectors made before.

```{r, fig.asp=0.8, fig.cap="OA shares of German research institutions per sector. The color of the boxes corresponds to the total publication output of the sector, their widths to the number of institutions per sector. Gray points display the OA shares for individual institutions."}
pubs_cat %>%
  group_by(sector) %>% 
  mutate(n_sector = n_distinct(PK_ITEMS),
         sector_cat = case_when(
           n_sector > 500000 ~ 'high_n_pubs',
           n_sector > 50000 ~ 'middle_n_pubs',
           TRUE ~ 'low_n_pubs'
         )) %>% 
  ungroup() %>% 
  left_join(oa_shares, by = 'INST_NAME') %>%
  group_by(INST_NAME, sector, oa_share, sector_cat) %>% 
  summarise() %>% 
  ggplot(aes(x = sector, y = oa_share)) + 
    geom_boxplot(aes(color = sector_cat), varwidth = TRUE, size = 1) +
    geom_jitter(alpha = 0.1) +
    scale_color_viridis(end = 0.8, labels = c("high", "medium", "low"), discrete = TRUE, direction = -1, option = "plasma")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "OA shares of institutions per sector", x = "", y = "OA share (in %)", color = "Publication output") 
```

### Prevalences of OA categories

- oa categories on the national level (relative numbers); faceted / dodged / lines ?
- oa vs. green (or other type): scatterplots
  - 1 point per sector (size of point according to pub output?)
  - faceted with all institutions per sector (facet grid? -> per sector and oa_category)
  
  
As mentioned in the previous chapters, there are several ways of providing open access to scientific journal articles. In this section, we want to investigate the prevalence of the most widespread OA routes: Green OA and Gold OA. We further distinguish these two main categories as described in the methodology section (see Table 1) according to whether the journal is fully OA (Gold OA), and into deposition on disciplinary, institutional, or OpenDOAR-listed repositories (Green OA). First, we visualise the number of articles per category on the national level, that is, not differentiated by sector. Note that, as mentioned before, the OA categories are not exclusive, that is, an article might be counted for several categories and numbers not necessarily sum up to the total number of articles published.

```{r, fig.asp=0.5, fig.cap="Development of the total number of journal articles over time per OA category. Categories are non-exclusive."}
cols_vir <- c(viridis(2, begin = 0.95, end = 1, direction = -1), viridis(4, begin = 0.3, end = 0.8, direction = -1), viridis(1, begin = 0, end = 0))
pubs_cat %>% 
  group_by(PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_line(aes(color = oa_category), size = 1) +
    scale_color_manual(values = cols_vir, name= "OA category") +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of Articles per OA category (national level)") +
    theme(legend.position = "right")
```
Observations:

- drop in disciplinary repos, maybe because of green embargoes / post-publish self-archiving
- drop in other oa journal -> Delayed OA
- apart from this: all OA categories increase, not oa decreases
- most prevalent category: other repo, i.e. not further specified green OA, followed closely by disciplinary


Or as barplot?

```{r, fig.asp=0.5, fig.cap="Development of the total number of journal articles over time per OA category. Categories are non-exclusive."}
cols_vir <- c(viridis(2, begin = 0.95, end = 1, direction = -1), viridis(4, begin = 0.3, end = 0.8, direction = -1), viridis(1, begin = 0, end = 0))
pubs_cat %>% 
  group_by(PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_col(aes(fill = oa_category), position = "dodge", color = "#b3b3b3a0", size = 0.2) +
    scale_fill_manual(values = cols_vir, name= "OA category") +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of Articles per OA category (national level)") +
    theme(legend.position = "right")
```

Faceted?


```{r, fig.asp=1, fig.cap="Development of the number of journal articles per OA category in comparison to the total number of journal articles published by the considered institutions over time. Categories are non-exclusive."}
all_articles <- pubs_cat%>% 
  group_by(PUBYEAR) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS))

cols_vir <- c(viridis(2, begin = 0.95, end = 1, direction = -1), viridis(4, begin = 0.3, end = 0.8, direction = -1), viridis(1, begin = 0, end = 0))
pubs_cat %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  group_by(PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_col(data = all_articles, aes(color = "All Articles"), fill = "#b3b3b3", size = 0.05) +
    scale_color_manual(values = "#ffffff", name = "", labels = "All Articles") +
    geom_col(aes(fill = oa_category)) +
    scale_fill_manual(values = cols_vir, name = "OA categories")+
    facet_wrap(~ oa_category, ncol = 2) +
    theme(strip.text.x = element_blank()) +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of Articles per OA host type (national level)") +
    theme(legend.position = "right")
```

Again, we go one step further and look at insitution specific OA proportions. In a first step, we calculate the category-specific OA shares of each institution.

```{r}
oa_shares_host_sec <- pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo")
  )) %>%
  group_by(sector) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>%
  group_by(sector, oa_category, n_total) %>% 
  summarise(n_cat = n_distinct(PK_ITEMS)) %>% 
  pivot_wider(names_from = oa_category,
              values_from = n_cat,
              values_fill = list(n_cat = 0)) %>% 
  mutate(oa_share_journal = journal_oa/n_total *100,
         oa_share_repo = repo_oa/n_total *100,
         oa_share_closed = not_oa/n_total * 100) %>%
  rename_at(vars(ends_with("_oa")), ~ paste0("n_", .)) %>% 
  ungroup()

oa_shares_cat <- pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  group_by(INST_NAME) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>%
  group_by(INST_NAME, oa_category, n_total) %>% 
  summarise(n_cat = n_distinct(PK_ITEMS)) %>% 
  pivot_wider(names_from = oa_category,
              values_from = n_cat,
              values_fill = list(n_cat = 0)) %>%
  mutate_at(c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"), list(share = ~ ./n_total*100)) %>% 
  rename_at(c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"), ~paste0("n_",.)) %>% 
  ungroup()
```

```{r, fig.asp=0.5, fig.cap="OA shares per sector. Shown is the relationship between Green OA shares and total OA shares. Point sizes correspond to total publication output."}
pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  )) %>% 
  group_by(sector) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>%
  group_by(sector, oa_category, n_total) %>% 
  mutate(n_cat = n_distinct(PK_ITEMS),
         share_cat = n_cat/n_total * 100) %>%
  ungroup() %>% 
  group_by(sector, oa_category, n_total, share_cat) %>% 
  summarise() %>% 
  pivot_wider(names_from = oa_category,
              values_from = share_cat,
              values_fill = list(share_cat = 0)) %>%
  pivot_longer(cols = c("journal_oa", "repo_oa"), names_to = "oa_route", values_to = "share") %>% 
  ggplot(aes(x = (100-not_oa), y = share)) +
    geom_point(aes(color = sector, size = n_total)) + 
    scale_color_viridis(end = 0.95, direction = -1, option = "plasma", discrete = TRUE, name = "Sectors")+
    guides(size = "none") +
    facet_wrap(~oa_route, labeller = labeller(oa_route = c("journal_oa" = "OA (publisher)", "repo_oa" = "OA (repository)"))) +
    labs(x = "OA share (in %)", y = "OA share (in %)", title = "Open Access shares of research sectors in Germany") +
    theme(legend.position = "right")
```

```{r, fig.asp=0.8, fig.cap="OA shares per sector. Shown is the relationship between OA shares of individual OA categories and overall OA share. Point sizes correspond to total publication output."}
pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  group_by(sector) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>%
  group_by(sector, oa_category, n_total) %>% 
  mutate(n_cat = n_distinct(PK_ITEMS),
         share_cat = n_cat/n_total * 100) %>%
  ungroup() %>% 
  group_by(sector, oa_category, n_total, share_cat) %>% 
  summarise() %>% 
  pivot_wider(names_from = oa_category,
              values_from = share_cat,
              values_fill = list(share_cat = 0)) %>%
  pivot_longer(cols = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"), names_to = "oa_route", values_to = "share") %>% 
  mutate(oa_route = factor(oa_route, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"))) %>% 
  ggplot(aes(x = (100-not_oa), y = share)) +
    geom_point(aes(color = sector, size = n_total)) + 
    scale_color_viridis(end = 0.95, direction = -1, option = "plasma", discrete = TRUE, name = "Sectors")+
    guides(size = "none") +
    facet_wrap(~oa_route, ncol =2) +
    labs(x = "OA share (in %)", y = "OA share (in %)", title = "Open Access shares of research sectors in Germany") +
    theme(legend.position = "right")
```
facet this per sector? plot oa_cat_shares for single sector, order by total oa_share (state in facet titles); similar to above bar/line plots faceted by sector (with/without yearly development)

Institution specific:

```{r, fig.asp=1, fig.cap="OA shares per OA category for individual research institutions. The relationship between OA shares of individual OA categories and overall OA share is shown."}
pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  group_by(INST_NAME) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>%
  group_by(sector) %>% 
  mutate(n_total_sec = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  group_by(INST_NAME, oa_category, n_total) %>% 
  mutate(n_cat = n_distinct(PK_ITEMS),
         share_cat = n_cat/n_total * 100) %>%
  ungroup() %>% 
  # group_by(INST_NAME) %>% 
  # mutate(inst_cat = case_when(
  #          n_total > 30000 ~ 'high_n_pubs',
  #          n_total > 6000 ~ 'middle_n_pubs',
  #          TRUE ~ 'low_n_pubs'
  #        )) %>% 
  # ungroup() %>% 
#  group_by(INST_NAME, sector, oa_category, n_total, share_cat, inst_cat) %>% 
  group_by(INST_NAME, sector, oa_category, n_total, n_total_sec, share_cat) %>% 
  summarise() %>% 
  pivot_wider(names_from = oa_category,
              values_from = share_cat,
              values_fill = list(share_cat = 0)) %>%
  pivot_longer(cols = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"), names_to = "oa_route", values_to = "share") %>% 
  mutate(oa_route = factor(oa_route, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"))) %>% 
  ggplot(aes(x = (100-not_oa), y = share)) +
#    geom_point(aes(color = inst_cat)) + 
#    scale_color_viridis(end = 0.95, direction = -1, labels = c("high", "medium", "low"), discrete = TRUE, option = "plasma", name = "Publication\n volume")+
    geom_point(aes(color = oa_route)) +
    scale_color_manual(values = cols_vir, name = "OA category")+
    facet_wrap(~oa_route, ncol = 2) +
      labs(x = "OA share (in %)", y = "OA share (in %)", title = "Open Access shares of research institutions in Germany") +
    theme(legend.position = "right")
```

```{r, fig.asp=1, fig.cap="OA shares per sector and OA category for individual institutions. The relationship between OA shares of individual OA categories and overall OA share is shown."}
pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  group_by(INST_NAME) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>%
  group_by(sector) %>% 
  mutate(n_total_sec = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  group_by(INST_NAME, oa_category, n_total) %>% 
  mutate(n_cat = n_distinct(PK_ITEMS),
         share_cat = n_cat/n_total * 100) %>%
  ungroup() %>% 
  # group_by(INST_NAME) %>% 
  # mutate(inst_cat = case_when(
  #          n_total > 30000 ~ 'high_n_pubs',
  #          n_total > 6000 ~ 'middle_n_pubs',
  #          TRUE ~ 'low_n_pubs'
  #        )) %>% 
  # ungroup() %>% 
#  group_by(INST_NAME, sector, oa_category, n_total, share_cat, inst_cat) %>% 
  group_by(INST_NAME, sector, oa_category, n_total, n_total_sec, share_cat) %>% 
  summarise() %>% 
  pivot_wider(names_from = oa_category,
              values_from = share_cat,
              values_fill = list(share_cat = 0)) %>%
  pivot_longer(cols = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"), names_to = "oa_route", values_to = "share") %>% 
  mutate(oa_route = factor(oa_route, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"))) %>% 
  ggplot(aes(x = (100-not_oa), y = share)) +
#    geom_point(aes(color = inst_cat)) + 
#    scale_color_viridis(end = 0.95, direction = -1, labels = c("high", "medium", "low"), discrete = TRUE, option = "plasma", name = "Publication\n volume")+
    geom_point(aes(color = oa_route)) +
    scale_color_manual(values = cols_vir, name = "OA category")+
#    facet_wrap(~oa_route, ncol = 2) +
    facet_grid(fct_rev(fct_reorder(sector, n_total_sec)) ~ oa_route) +
      labs(x = "OA share (in %)", y = "OA share (in %)", title = "Open Access shares of research institutions in Germany") +
    theme(legend.position = "bottom")
```
  
### Discussion

 - Upset plot of overlapping evidence categories to show influence of semantic scholar, webscraping.

In order to demonstrate the prevalence of evidence categories in Unpaywall, we load the original, non-categorized Unpaywall data:

```{r}
upw_evidence <- readr::read_csv("data/upw_evidence.csv")
upw_ev_cat <- upw_evidence %>% 
  mutate(upw_matched = ifelse(is.na(upw_doi), FALSE, TRUE)) %>% 
  filter(is_paratext == FALSE | is.na(is_paratext))
  rm(upw_evidence)
```

We now determine the evidence combinations for all matched DOIs and then calculate the frequency of each combination found.

```{r}
upw_ev_cat <- upw_ev_cat %>% 
  filter(upw_matched == TRUE) %>%
  select(upw_doi, evidence) %>% 
  arrange(upw_doi, evidence) %>% 
  aggregate(evidence ~ upw_doi, data = ., paste, collapse = "&")
upw_ev_cat_n <- upw_ev_cat %>% 
  group_by(evidence) %>% 
  summarise(n = n_distinct(upw_doi))
upw_ev_cat_n %>% 
  arrange(desc(n))
```
We now prepare the data for plotting with the UpSetR package and visualise the overlapping evidence categories.

```{r}
library(UpSetR)
# list with counts
upw_ev_upset_list <- as.list(upw_ev_cat_n$n)
# categories as list names
names(upw_ev_upset_list) <- upw_ev_cat_n$evidence
# convert to vector
evidence_categories_upset_expr <- unlist(upw_ev_upset_list)
upset(fromExpression(evidence_categories_upset_expr), nsets = 7, nintersects = 15, order.by = "freq", show.numbers = FALSE, set_size.angles = 25)

```


## Exemplary analytics on national, sectoral and institutional level

### National level

First, we look at the highest aggregation level, namely the national level. Here, publications of all institutions of the selected sectors (5+2) are counted per year and category. 

#### by host type

The first figure shows the development over time of the total number of publications per host type, that is, we collate all articles where access is provided via a publisher based platform as `OA (publisher)`, all repository based OA is collated as `OA (repository)`, and all remaining articles are labelled as `Not OA`.

```{r}
pubs_cat %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo")
  )) %>% 
  group_by(PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    #geom_line(aes(color = oa_category)) +
    geom_col(aes(fill = oa_category), position = "dodge") +
    # geom_bar(aes(fill = oa_category), stat = "identity", position = "dodge") +
    #scale_color_viridis(end = 0.85, discrete = TRUE, direction = -1) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA Categories", labels = c("OA (publisher)", "OA (repository)", "Not OA")) +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of Articles per OA host type (national level)") +
    theme(legend.position = "bottom")
```

In comparison to all articles published:

```{r}
all_articles <- pubs_cat%>% 
  group_by(PUBYEAR) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS))

pubs_cat %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  ))  %>% 
  group_by(PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
#  .$oa_category %>% levels()
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_col(data = all_articles, aes(color = "All Articles"), fill = "#b3b3b3", size = 0.05) +
    scale_color_manual(values = "#ffffff", name = "", labels = "All Articles") +
    geom_col(aes(fill = oa_category)) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "", labels = c("OA (publisher)", "OA (repository)", "Not OA"))+
    facet_wrap(~ oa_category, nrow = 1) +
    theme(strip.text.x = element_blank()) +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of Articles per OA host type (national level)") +
    theme(legend.position = "bottom")
```


Looking at proportions:

```{r}
pubs_cat %>% 
  group_by(PUBYEAR) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo")
  )) %>% 
  group_by(PUBYEAR, oa_category, n_total) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  mutate(prop = number_of_articles / n_total * 100) %>% 
  ggplot(aes(x = PUBYEAR, y = prop)) +
    #geom_line(aes(color = oa_category)) +
    geom_col(aes(fill = oa_category), position = "dodge") +
    # geom_bar(aes(fill = oa_category), stat = "identity", position = "dodge") +
    #scale_color_viridis(end = 0.85, discrete = TRUE, direction = -1) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, labels = c("OA (publisher)", "OA (repository)", "Not OA"), direction = -1) +
    labs(x = "Publication Year", y = "Proportion of Articles (in %)", title = "Proportion of Articles per OA host type", fill = "OA Category") +
    theme(legend.position = "right") 
    # theme(axis.title = element_text(size = 16),
    #     plot.title = element_text(size = 24),
    #     axis.text = element_text(size = 14),
    #     legend.text = element_text(size = 14),
    #     legend.title = element_text(size = 16))
```

As faceted graph:

```{r}
pubs_cat %>% 
  group_by(PUBYEAR) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo")
  )) %>% 
  group_by(PUBYEAR, oa_category, n_total) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  mutate(prop = number_of_articles / n_total * 100) %>% 
  ggplot(aes(x = PUBYEAR, y = prop)) +
    #geom_line(aes(color = oa_category)) +
    geom_col(aes(fill = oa_category)) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, labels = c("OA (publisher)", "OA (repository)", "Not OA"), direction = -1) +
    facet_wrap(~ oa_category, nrow = 1) +
    theme(strip.text.x = element_blank()) +
    labs(x = "Publication Year", y = "Proportion of Articles (in %)", title = "Proportion of Articles per OA host type", fill = "OA Category") +
    theme(legend.position = "bottom")
```

#### by OA category

The following figure shows very similar data as the one before - only that here, all OA categories from the schema are shown individually, that is, publisher based and repository based OA are further distinguished into subgroups.

```{r}
pubs_cat %>% 
  group_by(PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_line(aes(color = oa_category), size = 1) +
    #geom_col(aes(fill = oa_category), position = "dodge") +
    # geom_bar(aes(fill = oa_category), stat = "identity", position = "dodge") +
    #scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1) +
    scale_color_viridis(end = 0.95, discrete = TRUE, direction = -1, name= "OA category") +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of Articles per OA category (national level)") +
    theme(legend.position = "right")
```

In comparison to all articles published:

```{r}
all_articles <- pubs_cat%>% 
  group_by(PUBYEAR) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS))

pubs_cat %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  group_by(PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
#  .$oa_category %>% levels()
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_col(data = all_articles, aes(color = "All Articles"), fill = "#b3b3b3", size = 0.05) +
    scale_color_manual(values = "#ffffff", name = "", labels = "All Articles") +
    geom_col(aes(fill = oa_category)) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA categories")+
    facet_wrap(~ oa_category, ncol = 2) +
    theme(strip.text.x = element_blank()) +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of Articles per OA host type (national level)") +
    theme(legend.position = "right")
```

Looking at proportions, we get:

```{r}
pubs_cat %>% 
  group_by(PUBYEAR) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  group_by(PUBYEAR, oa_category, n_total) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
  mutate(prop = number_of_articles / n_total * 100) %>% 
  ggplot(aes(x = PUBYEAR, y = prop)) +
    #geom_line(aes(color = oa_category), size = 1) +
    geom_col(aes(fill = oa_category), position = "dodge") +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA category") +
    labs(x = "Publication Year", y = "Proportion of Articles (in %)", title = "Proportion of Articles per OA category (national level)") +
    theme(legend.position = "bottom")
```

As faceted graph:

```{r}
pubs_cat %>% 
  group_by(PUBYEAR) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  group_by(PUBYEAR, oa_category, n_total) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  mutate(prop = number_of_articles / n_total * 100) %>% 
  ggplot(aes(x = PUBYEAR, y = prop)) +
    #geom_line(aes(color = oa_category)) +
    geom_col(aes(fill = oa_category)) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA category") +
    facet_wrap(~ oa_category, ncol = 2) +
    theme(strip.text.x = element_blank()) +
    labs(x = "Publication Year", y = "Proportion of Articles (in %)", title = "Proportion of Articles per OA host type", fill = "OA Category") +
    theme(legend.position = "right")
```

### Level of sectors

Next, we go one level down and look at how the separate sectors developed.

Number of OA articles per sector in comparison to all articles published within the sector:

```{r, fig.asp=1}
pubs_cat %>% 
  mutate(oa_category = fct_relevel(factor(oa_category), "full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa")) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    is_oa = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  ))  %>%
  group_by(sector, PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>% 
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_area(aes(fill = fct_rev(oa_category), group = fct_rev(oa_category)),  alpha = 0.8) +
    scale_fill_manual(values = c("#b3b3b3a0", "#56B4E9"), name = "Is OA?", labels = c("Not OA", "OA"))+
    facet_wrap(~ sector, scales = "free", ncol = 2) +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of OA Articles (level of sectors)") +
    theme(legend.position = "right")
```

#### by host type

The first figure depicts the number of publications per host type category over time for each sector.

```{r, fig.asp=1}
#, fig.asp=1.5}
pubs_cat %>% 
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  )) %>% 
  group_by(sector, PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_col(aes(fill = oa_category), position = "dodge") +
    facet_wrap(~ sector, scales = "free", ncol = 2) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA category", labels = c("OA (publisher)", "OA (repository)", "Not OA")) +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of Articles per OA host type (level of sectors)", color = "oa_category") +
    theme(legend.position = "right")
```

Looking at proportions:
```{r, fig.asp=1}
#, fig.asp=1.5}
pubs_cat %>%
  group_by(PUBYEAR, sector) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  )) %>% 
  group_by(sector, PUBYEAR, oa_category, n_total) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  mutate(prop = number_of_articles / n_total *100) %>% 
  ggplot(aes(x = PUBYEAR, y = prop)) +
    geom_col(aes(fill = oa_category), position = "dodge") +
    facet_wrap(~ sector, scales = "free") +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA category", labels = c("OA (publisher)", "OA (repository)", "Not OA")) +
    labs(x = "Publication Year", y = "Proportion of Articles (in %)", title = "Proportion of Articles per OA host type (level of sectors)", color = "oa_category") +
    theme(legend.position = "bottom")
```

Facet grid:

```{r, fig.asp=1}
pubs_cat_abbr <- pubs_cat %>% 
  mutate(sector = case_when(
    sector == "Hochschulen" ~ "HS",
    sector == "Leibniz-Gemeinschaft" ~ "LG",
    sector == "Helmholtz-Gemeinschaft" ~ "HG",
    sector == "Max-Planck-Gesellschaft" ~ "MPG",
    sector == "Ressortforschung-Bund" ~ "R-B",
    sector == "Ressortforschung-Länder" ~ "R-L",
    sector == "Fraunhofer-Gesellschaft" ~ "FG"
  ))
pubs_cat_abbr %>%
  group_by(PUBYEAR, sector) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>%
    mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  )) %>% 
  group_by(sector, PUBYEAR, oa_category, n_total) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  mutate(prop = number_of_articles / n_total * 100) %>% 
  ggplot(aes(x = PUBYEAR, y = prop)) +
    geom_col(aes(fill = oa_category)) +
    facet_grid(sector ~ oa_category) +
    theme(strip.text.x = element_blank()) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA category", labels = c("OA (publisher)", "OA (repository)", "Not OA")) +
    labs(x = "Publication Year", y = "Proportion of Articles (in %)", title = "Proportion of Articles per OA host type (level of sectors)", color = "oa_category") +
    theme(legend.position = "bottom")
```

#### by OA category

Distinguishing all the single OA categories:

```{r, fig.asp=1}
pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  group_by(sector, PUBYEAR, oa_category) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  ggplot(aes(x = PUBYEAR, y = number_of_articles)) +
    geom_col(aes(fill = oa_category), position = "dodge") +
    facet_wrap(~ sector, scales = "free", ncol = 2) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA category") +
    labs(x = "Publication Year", y = "Number of Articles", title = "Number of Articles per OA category (level of sectors)", color = "oa_category") +
    theme(legend.position = "bottom")
```

Looking at proportions:

```{r, fig.asp=1}
pubs_cat %>%
  group_by(PUBYEAR, sector) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  group_by(sector, PUBYEAR, oa_category, n_total) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  mutate(prop = number_of_articles / n_total * 100) %>% 
  ggplot(aes(x = PUBYEAR, y = prop)) +
    geom_col(aes(fill = oa_category), position = "dodge") +
    facet_wrap(~ sector, scales = "free", ncol = 2) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA category") +
    labs(x = "Publication Year", y = "Proportion of Articles (in %)", title = "Proportion of Articles per OA category (level of sectors)", color = "oa_category") +
    theme(legend.position = "bottom")
```


Facet grid:

```{r, fig.asp=1.2}
pubs_cat_abbr <- pubs_cat %>% 
  mutate(sector = case_when(
    sector == "Hochschulen" ~ "HS",
    sector == "Leibniz-Gemeinschaft" ~ "LG",
    sector == "Helmholtz-Gemeinschaft" ~ "HG",
    sector == "Max-Planck-Gesellschaft" ~ "MPG",
    sector == "Ressortforschung-Bund" ~ "R-B",
    sector == "Ressortforschung-Länder" ~ "R-L",
    sector == "Fraunhofer-Gesellschaft" ~ "FG"
  ))
pubs_cat_abbr %>%
  group_by(PUBYEAR, sector) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>% 
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>%
  group_by(sector, PUBYEAR, oa_category, n_total) %>% 
  summarise(number_of_articles = n_distinct(PK_ITEMS)) %>%
  mutate(prop = number_of_articles / n_total * 100) %>% 
  ggplot(aes(x = PUBYEAR, y = prop)) +
    geom_col(aes(fill = oa_category)) +
    facet_grid(sector ~ oa_category) +
    theme(strip.text.x = element_blank()) +
    scale_fill_viridis(end = 0.95, discrete = TRUE, direction = -1, name = "OA category") +
    labs(x = "Publication Year", y = "Proportion of Articles (in %)", title = "Proportion of Articles per OA category (level of sectors)", color = "oa_category") +
    theme(legend.position = "bottom")
```

### Institutional level

Lastly, we go down to the institutional level. Since there are more than 400 institutions in the dataset, we do not want to present figures for individual institutions. Instead, we want to show the variety of open access shares and strategies. To this end, we first calculate the individual OA shares.

```{r}
oa_shares <- pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    is_oa = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  )) %>% 
  group_by(INST_NAME) %>% 
  mutate(n_total = n_distinct(PK_ITEMS)) %>% 
  ungroup() %>%
  group_by(INST_NAME, oa_category, n_total) %>% 
  summarise(n_cat = n_distinct(PK_ITEMS)) %>% 
  pivot_wider(names_from = oa_category,
              values_from = n_cat,
              values_fill = list(n_cat = 0)) %>% 
  mutate(oa_share = is_oa/n_total *100) %>%
  rename(n_oa = is_oa, n_not_oa = not_oa) %>% 
  ungroup()
```

We now display the variability among institutions per sector with respect to the total publication output.

```{r, fig.asp=1}
pubs_cat %>%
  mutate(oa_category = factor(oa_category, levels = c("full_oa_journal", "other_oa_journal", "opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo", "not_oa"))) %>% 
  mutate(oa_category = fct_collapse(oa_category,
    journal_oa = c("full_oa_journal", "other_oa_journal"),
    repo_oa = c("opendoar_inst", "opendoar_subject", "opendoar_other", "other_repo"),
    not_oa = "not_oa"
  )) %>% 
  left_join(oa_shares, by = 'INST_NAME') %>% 
  group_by(sector, INST_NAME, n_total, oa_share) %>%
  summarise() %>% 
  ggplot(aes(x = n_total, y = oa_share)) +
    geom_point(aes(color = sector)) + 
    scale_x_log10() + 
    scale_color_viridis(end = 0.95, direction = -1, option = "plasma", discrete = TRUE, name = "Sectors")+
    facet_wrap(~ sector, ncol = 2) +
    labs(x = "Total publication output (logarithmic scale)", y = "OA share (in %)", title = "Open Access shares of research institutions in Germany") +
    theme(legend.position = "right")
```

Sectorwise comparison of OA shares

```{r, fig.asp=0.8}
pubs_cat %>%
  left_join(oa_shares, by = 'INST_NAME') %>%
  group_by(INST_NAME, sector, oa_share, n_total) %>% 
  summarise() %>% 
  group_by(sector) %>% 
  summarise_at(vars(n_total, oa_share), funs(min, max, mean, median,sd)) %>% 
  select(sector, starts_with('oa_share'), starts_with('n_total'))

n_fun <- function(x){
  return(data.frame(y = 105,
                    label = paste ("n=", length(x))))
}

pubs_cat %>%
  group_by(sector) %>% 
  mutate(n_sector = n_distinct(PK_ITEMS),
         sector_cat = case_when(
           n_sector > 500000 ~ 'high_n_pubs',
           n_sector > 50000 ~ 'middle_n_pubs',
           TRUE ~ 'low_n_pubs'
         )) %>% 
  ungroup() %>% 
  left_join(oa_shares, by = 'INST_NAME') %>%
  group_by(INST_NAME, sector, oa_share, sector_cat) %>% 
  summarise() %>% 
  ggplot(aes(x = sector, y = oa_share)) + 
    geom_boxplot(aes(color = sector_cat), varwidth = TRUE, size = 1) +
#    geom_boxplot() +
 #   stat_summary(fun.data = n_fun, geom = "text", hjust = 0.5) +
    geom_jitter(alpha = 0.1) +
    scale_color_viridis(end = 0.8, labels = c("high", "medium", "low"), discrete = TRUE, direction = -1, option = "plasma")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "OA shares of institutions per sector", x = "", y = "OA share (in %)", color = "Publication output",
         caption = "Widths of the boxes corresponds to the number of institutions per sector,\n gray points display oa shares for individual institutions.") 
      # theme(axis.title = element_text(size = 24),
      #   plot.title = element_text(size = 36),
      #   axis.text = element_text(size = 21),
      #   legend.text = element_text(size = 21),
      #   legend.title = element_text(size = 24))
```
