---
title: "Some preliminary data exploration"
author: "Anne Hobert"
output: github_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
```

```{r}
library(tidyverse)
library(urltools)
library(cowplot)
```


Here, we perform some preliminary exploration of the generated datasets. We obtain tibbles `pubs_wos`, `upw_evidence`, the matched data frame `pubs_wos_upw`, and the final dataframe `pubs_cat` as described in [data_gathering.Rmd](data_gathering.Rmd).

## How many publications are there? How many can be matched to Unpaywall?

There are `r pubs_wos %>% summarise(n = n_distinct(PK_ITEMS)) %>% .$n` different articles in our sample (differentiated by pk_items). `r pubs_wos %>% filter(is.na(DOI)) %>% summarise(n = n_distinct(PK_ITEMS)) %>% .$n` of them do not have any DOI information. `r pubs_wos %>% group_by(DOI) %>%   summarise(n = n_distinct(PK_ITEMS)) %>% filter(n==2) %>% count() %>% .$n` DOIs appear twice, that is, they are associated with two different `PK_ITEMS`. The remaining DOIs are linked to a unique `PK_ITEMS`. In total, there are `r pubs_wos %>% filter(!is.na(DOI)) %>% summarise(n=n_distinct(DOI)) %>% .$n` distinct DOIs within the sample. Of these, `r upw_evidence %>% filter(!is.na(upw_doi)) %>% summarise(n = n_distinct(wos_doi)) %>% .$n` are matched to Unpaywall records, whereas `r upw_evidence %>% filter(is.na(upw_doi)) %>% summarise(n = n_distinct(wos_doi)) %>% .$n`, or `r round(upw_evidence %>% filter(is.na(upw_doi)) %>% summarise(n = n_distinct(wos_doi)) %>% .$n/(upw_evidence %>% summarise(n = n_distinct(wos_doi)) %>% .$n), 2)*100` %, did not correspond to any of the DOIs in our Unpaywall dataset.


Briefly evaluate the matching procedure of matching WoS and Unpaywall: The following table shows the number of articles from WoS that have no DOI information (`no_doi`), articles that have DOI information in WoS but could not be matched to any record in the Unpwaywall dataset we used (`unmatched`), articles matched to Unpaywall for which Unpaywall did not find any open version (`upw_closed`) and articles with OA evidence in Unpaywall (`OA`).
```{r}
pubs_wos_upw %>% 
  mutate(matched_oa = case_when(
    is.na(DOI) ~ 'no_doi',
    is.na(upw_doi) ~ 'unmatched',
    !is_oa ~ 'upw_closed',
    TRUE ~ 'OA'
  )) %>% 
  group_by(matched_oa) %>% 
  summarise(n = n_distinct(PK_ITEMS)) %>% 
  # filter(!matched_oa %in% c('unmatched', 'no_doi')) %>% 
  mutate(prop = n / sum(n))%>% 
  knitr::kable()
```

## Number of Full-Text links per OA Article 
```{r}
oa_fxt_count <- pubs_wos_upw %>% 
  filter(!is.na(host_type)) %>%
  group_by(DOI) %>%
  mutate(n_oa_versions = n()) %>%
  ungroup() %>%
  group_by(DOI, n_oa_versions, host_type) %>%
  summarise(ftxt_n = n()) %>%
  ungroup()
```

```{r}
oa_fxt_count %>%
  distinct(DOI, n_oa_versions) %>%
  ggplot(aes(x = n_oa_versions)) +
  geom_histogram(binwidth = 1, fill="#56B4E9", alpha=0.5, color = "#56B4E9") +
  geom_vline(aes(xintercept = mean(n_oa_versions, na.rm = T)),
             colour = "#E69F00", linetype ="dashed", size = .8) +
  geom_vline(aes(xintercept = median(n_oa_versions, na.rm = T)),
             colour = "red", linetype ="dashed", size = .8) +
  theme_minimal_hgrid(12) +
  labs(x = "Number of Full-Text Links",
       y = "German OA Articles")
```

### per host type

```{r}
# compute densities for full text numbers lengths
oa_fxt_count %>%
  #filter(!is.na(host_type)) %>% 
  #distinct() %>% 
  ggplot(aes(x = ftxt_n, fill = host_type, color = host_type)) +
  geom_histogram(binwidth = 1, alpha=0.5) +
  facet_grid(~ host_type, scales = "free_x") +
  theme_minimal_hgrid(12) +
  labs(x = "Number of Full-Text Links",
       y = "German OA Articles") 
```

## Number of OA Articles per Host Type

```{r}
oa_fxt_count %>% 
  group_by(DOI) %>%
  mutate(n_host_type = n()) %>%
  mutate(host_type = ifelse(n_host_type == 2, "publisher & repository", host_type)) %>%
  ungroup() %>%
  group_by(host_type) %>%
  summarise(n = n_distinct(DOI)) %>%
  mutate(prop = n / sum(n)) %>%
  knitr::kable()
```

### Repository analysis using OpenDOAR

```{r}
 repo_df <- pubs_wos_upw %>% 
  distinct() %>%
  filter(host_type == "repository") %>%
  mutate(url_domain = domain(url_for_pdf)) %>%
  mutate(url_domain = gsub("www.", "", url_domain)) %>%
  mutate(landing_domain = domain(url_for_landing_page)) %>%
  mutate(landing_domain = gsub("www.", "", url_domain))
```

```{r}
opendoar <- readr::read_csv("data/opendoar_data_tidier.csv") %>%
  mutate(repo_domain = domain(repo_url)) %>%
  mutate(repo_domain = gsub("www.", "", repo_domain))
```


#### Number of self-archived full-texts by repository type

```{r}
repo_df %>% 
  distinct(DOI, url_domain, landing_domain) %>% 
  left_join(opendoar,  by = c("url_domain" = "repo_domain", "landing_domain" = "repo_domain")) %>% 
  ungroup() %>% 
  group_by(repo_type) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  mutate(prop = n / sum(n)) %>%
  knitr::kable()
```

without "pdfs.semanticscholar.org".

```{r}
repo_df %>% 
  distinct(DOI, url_domain, landing_domain) %>% 
  filter(url_domain != "pdfs.semanticscholar.org") %>%
  left_join(opendoar,  by = c("url_domain" = "repo_domain", "landing_domain" = "repo_domain")) %>% 
  ungroup() %>% 
  group_by(repo_type) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  mutate(prop = n / sum(n)) %>%
  knitr::kable()
```

### Top Repository Domains

```{r}
repo_df %>%
  group_by(url_domain) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n) * 100) %>%
  arrange(desc(n)) %>%
  head(20) %>%
  knitr::kable()
```

```{r}
repo_df %>% 
  distinct(url_domain, landing_domain) %>% 
  left_join(opendoar,  by = c("url_domain" = "repo_domain", "landing_domain" = "repo_domain"))

repo_df %>% 
  distinct(FK_ITEMS, DOI, url_domain, landing_domain) %>% 
  left_join(opendoar,  by = c("url_domain" = "repo_domain", "landing_domain" = "repo_domain")) %>% 
  ungroup() %>% 
  write_csv("data/german_unis_url_matching.csv")
```

## Institutional level

### Outliers with all OA publications

Investigate, which institutions are the outliers with only open access articles.

```{r}
outliers <- pubs_cat %>%
  filter(INST_NAME %in% c("Heidelberg Graduate School of Fundamental Physics", "Schloss Dagstuhl - Leibniz-Zentrum für Informatik GmbH (LZI)"))
outliers %>%
  group_by(INST_NAME) %>% 
  summarise(n = n_distinct(PK_ITEMS))
outliers %>% 
  group_by(INST_NAME, oa_category) %>% 
  summarise(n = n_distinct(PK_ITEMS)) %>% 
  arrange(INST_NAME, desc(n))
outliers %>%
  group_by(INST_NAME, url_domain, oa_category) %>%
  summarise(n = n_distinct(PK_ITEMS)) %>% 
  arrange(desc(n))
```
Schloss Dagstuhl with just one publication, that is OA. 2nd one is Heidelberg Graduate School of Fundamental Physics with 56 publications, 55 of which are published in the journal Astronomy and Astrophysics. Most of these publications are tagged as publisher based OA but are also deposited in various repositories.
